Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-10 arxiv/tools.tex
sec C

\begin{lemma}[Lemma 3.8 of \cite{EKYY1} and Theorem B.1 of \cite{Delocal}]\label{largedeviation}
Let $(x_i)$, $(y_j)$ be independent families of centered and independent random variables, and $(A_i)$, $(B_{ij})$ be families of deterministic complex numbers. Suppose the entries $x_i$ and $y_j$ have variances at most $1$, and $n^{-1/2}x_i$ and $n^{-1/2}y_j$ satisfy the bounded support condition (\ref{eq_support}) with $q\le n^{-\phi}$ for a small constant $\phi>0$. Then we have the following bounds:
%for any fixed $\xi>0$, the following bounds hold with $\xi$-high probability:
\begin{align*}
\Big| \frac{1}{\sqrt{n}}\sum_i A_i x_i \Big\vert \prec  q \max_{i} \vert A_i \vert+ \frac{1}{\sqrt{n}}\Big(\sum_i |A_i|^2 \Big)^{1/2} , \quad &\Big\vert \frac1n\sum_{i,j} x_i B_{ij} y_j \Big\vert \prec q^2 B_d  + qB_o + \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}}, \\
\Big\vert \frac1n\sum_{i} (|x_i|^2-\mathbb E|x_i|^2) B_{ii}  \Big\vert  \prec q B_d   , \quad &\Big\vert \frac1n\sum_{i\ne j} \bar x_i B_{ij} x_j \Big\vert  \prec qB_o + \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} ,
\end{align*}
where $B_d:=\max_{i} |B_{ii} |$ and $B_o:= \max_{i\ne j} |B_{ij}|.$ Moreover, if all the moments of $ x_i$ and $ y_j$ exist in the sense of equation \eqref{assmAhigh2}, then we have stronger bounds
\begin{align}
\Big\vert \frac1{\sqrt{n}}\sum_i A_i x_i \Big\vert \prec  \frac{1}{\sqrt{n}}\Big(\sum_i |A_i|^2 \Big)^{1/2} , \quad  & \Big\vert \frac1n\sum_{i,j} x_i B_{ij} y_j \Big\vert \prec  \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} ,\label{eq largedev1} \\
 \Big\vert  \frac1n \sum_{i} (|x_i|^2-\mathbb E|x_i|^2) B_{ii}  \Big\vert  \prec \frac1{n}\Big( \sum_i |B_{ii} |^2\Big)^{1/2}  ,\quad & \Big\vert \frac1n\sum_{i\ne j} \bar x_i B_{ij} x_j \Big\vert  \prec  \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} .\label{eq largedev2}
\end{align}
\end{lemma}

>>>>

\begin{lemma}[Lemma 3.8 of \cite{EKYY1} and Theorem B.1 of \cite{Delocal}]\label{largedeviation}
Let $(x_i)$, $(y_j)$ be independent families of centered and independent random variables, and $(A_i)$, $(B_{ij})$ be families of deterministic complex numbers. Suppose the entries $x_i$ and $y_j$ have variances at most $1$, and $ x_i$ and $ y_j$ satisfy the bounded support condition (\ref{eq_support}) for a deterministic parameter $q$. %with $q\le n^{-\phi}$ for a small constant $\phi>0$. 
Then we have the following bounds:
%for any fixed $\xi>0$, the following bounds hold with $\xi$-high probability:
\begin{align}
\Big| \sum_{i=1}^n A_i x_i \Big\vert \prec  q \max_{i} \vert A_i \vert+ \Big(\sum_i |A_i|^2 \Big)^{1/2} , \quad &\Big\vert  \sum_{i,j=1}^n x_i B_{ij} y_j \Big\vert \prec q^2 B_d  + q n^{1/2}B_o +  \Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}},\label{eq largedev10}  \\
\Big\vert  \sum_{i=1}^n (|x_i|^2-\mathbb E|x_i|^2) B_{ii}  \Big\vert  \prec q n^{1/2}B_d   , \quad &\Big\vert  \sum_{1\le i\ne j\le n} \bar x_i B_{ij} x_j \Big\vert  \prec qn^{1/2}B_o +  \Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} ,\label{eq largedev20} 
\end{align}
where $B_d:=\max_{i} |B_{ii} |$ and $B_o:= \max_{i\ne j} |B_{ij}|.$ Moreover, if all the moments of $ x_i$ and $ y_j$ exist in the sense of equation \eqref{assmAhigh2}, then we have stronger bounds
\begin{align}
\Big\vert \sum_i A_i x_i \Big\vert \prec  \Big(\sum_i |A_i|^2 \Big)^{1/2} , \quad  & \Big\vert \sum_{i,j} x_i B_{ij} y_j \Big\vert \prec  \Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} ,\label{eq largedev1} \\
 \Big\vert  \sum_{i} (|x_i|^2-\mathbb E|x_i|^2) B_{ii}  \Big\vert  \prec  \Big( \sum_i |B_{ii} |^2\Big)^{1/2}  ,\quad & \Big\vert  \sum_{i\ne j} \bar x_i B_{ij} x_j \Big\vert  \prec \Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} .\label{eq largedev2}
\end{align}
%\begin{align}
%\Big\vert \frac1{\sqrt{n}}\sum_i A_i x_i \Big\vert \prec  \frac{1}{\sqrt{n}}\Big(\sum_i |A_i|^2 \Big)^{1/2} , \quad  & \Big\vert \frac1n\sum_{i,j} x_i B_{ij} y_j \Big\vert \prec  \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} ,\label{eq largedev1} \\
% \Big\vert  \frac1n \sum_{i} (|x_i|^2-\mathbb E|x_i|^2) B_{ii}  \Big\vert  \prec \frac1{n}\Big( \sum_i |B_{ii} |^2\Big)^{1/2}  ,\quad & \Big\vert \frac1n\sum_{i\ne j} \bar x_i B_{ij} x_j \Big\vert  \prec  \frac{1}{n}\Big(\sum_{i\ne j} |B_{ij}|^2\Big)^{{1}/{2}} .\label{eq largedev2}
%\end{align}
\end{lemma}

As a corollary, we can obtain the following concentration estimates for random variables with bounded moments as in equation \eqref{assmAhigh}.

\begin{corollary} \label{cor_largedeviation}
Let $Z\in \R^{n\times p}$ be a random matrix as in Assumption \ref{assume_rm}. Then for any deterministic vector $v\in \R^p$, we have that
$$\left|\|Zv\|^2- n\|v\|^2\right|\le  n^{1-2c_\varphi} \max_{1\le i\le p}|v(i)|^2$$
with high probability.
\end{corollary}
 \begin{proof}
 Using \eqref{eq largedev10} and \eqref{eq largedev20}
 \end{proof}
